---
title: "A Brief Introduction To R"
author: "Robert Olendorf"
date: "February 16, 2016"
output: html_document
---

This document is a brief description of using R Studio, covering the using the **Console**, data types, the __*Environment*__ and **_History Tabs_** at the top right, and the **_Plots_** and **_Packages_** Tabs in the lower right. This document was written in the document viewer in the top left. We'll create a markdown document at the same time. From the **[File]** menu select **[New File]** option and choose **[R Script]**. 

First ```1 + 1``` into the Console. Now insert it into your document using the history.  Make sure your new script is active on the top left and go to your **_History_** tab choose the line _1+1_ and click **[To Source]**. It will insert the line into your document. This is a great workflow in R. Try stuff out on the console, then when you have what you want, use your history to insert it into a script to save it. That is one of the basics of reproducible research. FYI, you can also select and insert multiple lines.


## Variables and Datatypes

So lets explore R a bit. First variables and data types. We can assign a value to a variable like so. Type the following in your console. I'll be adding it to my hello_r document as well. 

```{r}
  a <- 1
  a
```

R is dynamically typed. So it knows (or more correctly assumes) that a should be an integer. If we reassign it to other types, it will handle it easily.

```{r}
  # A float
  a <- 1.1
  a
  
  # A character
  a <- "a"
  a
  
  # Logical
  a <- TRUE
  a
```
Identifiers consist of a sequence of letters, digits, the period (‘.’) and the underscore. They must not start with a digit or an underscore, or with a period followed by a digit.

## Data Structures

R has a few data structures too. 

### Vectors

Vectors are essentially an array. They all have to be the same data type.

```{r}
  a <- c(1,2,3,4)  # numeric vector
  a 
  
  b <- c("one", "2", "three") # character vector
  b
  
  b.1 <- c(1, "2", 3) # R forces this to all characters
  b.1
  
  c <- c(TRUE, FALSE, TRUE, FALSE) #logical vector
  c
```

### Matrices


