---
title: "Hello R"
author: "Robert Olendorf"
date: "February 16, 2016"
output: html_document
---

## Hello R

Type ```1 + 1``` into the console and hit return. You should see the answer. 


```{r}
# This is a comment, and as simple as it gets.
1 + 1
```

Setting a variable. Notice I enter the **_a_** again to print it. Otherwise it would only display the assignment without a result. R is dynamically typed, so it makes an assumption about what you want.

```{r}
  # An integer
  a <- 1
  a
  
  # A float
  a <- 1.1
  a
  
  # A character
  a <- "b"
  a
  
  # If I want to have the character one, I have to quote it.
  a <- "1"
  a
  # Im using a built in function here.
  is.numeric(a)
  
  # Logical
  a <- TRUE
  a
  
```

Identifiers consist of a sequence of letters, digits, the period (‘.’) and the underscore. They must not start with a digit or an underscore, or with a period followed by a digit. Periods are more often used than underscores in R. They don't have a special meaning, but are often used to unofficially namespace things 

```{r}
  # This works
  ab1.my.things = 3
  ab1.my.things
  
  # This doesn't
  "1.my.things = 4"
```

## Data Structures

R has a few data structures too. 

### Vectors

Vectors are essentially an array. They all have to be the same data type.

```{r}
  a <- c(1,2,3,4)  # numeric vector
  a 
  
  b <- c("one", "2", "three") # character vector
  b
  
  b.1 <- c(1, "2", 3) # R forces this to all characters
  b.1
  
  c <- c(TRUE, FALSE, TRUE, FALSE) #logical vector
  c
```

You can also create vectors quickly using the following syntaxes.

```{r}
  a <- 0:5
  a
  
  b <- seq(0, 5, 0.5)
  b
```

### Matrices







